---
title: "Making State Tibbles To Support Graphical Comparisons"
author: "Brandon Tao"
date: "9/21/2020"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


### 0. Setup


```{r message=FALSE}
library(tidyverse)
```

State statistics are often available in csv files.  Here we
read previously produced Covid-19 state case and death count
csv files.  They have three September days to show short temporal
trends.  

### 1. Read csv files

Note that this project has Data and Docs directories to provide
more organized storage.  The data and documents are no
long mixed with R scripts in a single directory.  One small cost is
needing to specify the Data directory as shown in the script below. 

```{r message=FALSE}
state_pop <- read_csv('Data/State_Pop_2019.csv')
names(state_pop) <- c('State','Postal','Fips','Pop')
state_pop <- select(state_pop, Postal, Pop)

cases <- read_csv('Data/cases_Sep6_13_20.csv')
nam <- names(cases)
nam[3] <- 'Postal'
names(cases) <- nam

deaths <- read_csv('Data/deaths_Sep6_13_20.csv')
names(deaths) <- nam  # same as above
```

```{r}
cases <- cases %>% filter(
	countyFIPS != 2270 & countyFIPS != 6000)
deaths <- deaths %>% filter(
	countyFIPS != 2270 & countyFIPS != 6000)
	
```

### 2. Use group_by and summarize to get state counts for chosen days
  
  
```{r message=FALSE}

ST_cases <- cases %>% group_by(Postal)%>%
  summarize(Sep6=sum(Sep_06_2020),
            Sep13=sum(Sep_13_2020),   
            Sep20=sum(Sep_20_2020))


ST_deaths<- deaths %>% group_by(Postal)%>%
  summarize(Sep6=sum(Sep_06_2020),
            Sep13=sum(Sep_13_2020),   
            Sep20=sum(Sep_20_2020))

```

### 3. Use left_join and mutate to compute rates per 100000


```{r}

ST_case_rates  <- left_join(ST_cases,state_pop,by='Postal') %>% 
           mutate(Sep6r=100000*Sep6/Pop,
                  Sep13r=100000*Sep13/Pop,
                  Sep20r=100000*Sep20/Pop,
           type=rep('Cases',nrow(ST_cases)))

ST_death_rates  <- left_join(ST_deaths,state_pop,by='Postal') %>% 
           mutate(Sep6r=100000*Sep6/Pop,
                  Sep13r=100000*Sep13/Pop,
                  Sep20r=100000*Sep20/Pop,
           type=rep('Deaths',nrow(ST_cases)))

# A quick look
ST_case_rates
ST_death_rates

```

### 4. Save rate tibbles for use in graphics


Here we save the tibbles into an RData file.  This is easy to load into 
working memory.  Alternately we could write csv files and read them back in later.

```{r}
save(ST_case_rates,ST_death_rates,file='Data/ST_rates_Sep6_13_20.RData')
```

 



