---
title: "Turn Covid-19 Dates in R column names and Select variables"
author: "Brandon Tao"
date: "9/21/2020"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

The USAfacts Covid-19 case and death csv files provides
a good resource for class room examples, homework assignments
and possibly class projects.  

The csv files pose a challenge because the first row has
dates that are not allowed as R variable names. This script
produces dates that are valid column names and selects few dates
to provide smaller csv files for student use. 

The important class topics are the often used data wrangling functions
and graphics.  The need to modify dates is relatively
uncommon.  This script is shown in class and provided for
possible future reference.  Modification of dates is not a
quiz topic.


```{r}
library(tidyverse)
library(anytime)
```

### 1. Read the USAfacts US county cases and deaths csv files 

The US county files are in the Data directory of this project.

```{r}

cases  <- read_csv(
  file='Data/cases.csv')

deaths  <- read_csv(
  file='Data/deaths.csv')

#check <- names(cases)
# check

```

The first two variable names, countyFIPs and "County Name",
are county identifiers.  The next two names, state and stateFIPS,
are state identifiers. The remaining column names are dates
from 1/22/2020 to 9/20/2020.  
  
In R, variable names must start with a letter and have only
letters, numbers, an underscore, or a period. 

The County Name has a blank that can be omitted.
  
The dates start with a number. How can this be fixed?

####  2. Make valid variable names for data

We can fix date by putting a three letter month first
and by separating the month, day and year with two underscores. 
Here we use '_' but we could use  '.'.

```{r}

# Get the column names as vector
tmp <- names(cases)

# Fix the County Name in the first 4 names
first4 <- tmp[1:4] 
first4[2] <- 'CountyName'

# get the dates by removing first 4 names 
rest <- tmp[-(1:4)]
rest[1:2]  # look as the first two
 
# put numbers in the standard year month day order 
rest <- anydate(rest)
rest[1:2] 

# Change the format
# Below the symbol %b means use the 3 letter month name.
# The symbol %d means use the number.
# The symbol %Y means use the 4 digit number.
# the separators '_' could be replaced by '.'.

rest<- format(rest,"%b_%d_%Y") # Reformat '.'  
rest[1:2]
names(cases) <- c(first4,rest) # rename all columns

```
  
Now we fix the deaths column names.  Note, this script and
edited copy of the case script above.  It could be 
turned into a function.  All that differs is the
csv file name.  This file namee would the function argument.   
    
```{r}

deaths  <- read_csv(file='Data/deaths.csv')
tmp <- names(deaths)

# Fix the County Name in the first for names
first4 <- tmp[1:4] 
first4[2] <- 'CountyName'

# get the dates
rest <- tmp[-(1:4)]
 
# put numbers in the standard year month day order 
rest <- anydate(rest)

# Change the format
# Below the symbol %b means use 3 letter month name.
# The symbol %d means use the number
# The symbol %Y means use the 4 digit number
# the separators '_' could be replaced by '.'.
rest<- format(rest,"%b_%d_%Y") # Reformat '.'  
rest[1:2]
names(deaths) <- c(first4,rest)
```

### 3. Select first four variables and the last 3 days that are 7 days apart

```{r}

# names(cases)

cases_Sep6_13_20 <- select(cases, countyFIPS:stateFIPS,
                   Sep_06_2020, Sep_13_2020, Sep_20_2020)
names(cases_Sep6_13_20)

deaths_Sep6_13_20 <- select(deaths, countyFIPS:stateFIPS,
                    Sep_06_2020, Sep_13_2020, Sep_20_2020)


names(cases_Sep6_13_20)
names(deaths_Sep6_13_20)

View(cases_Sep6_13_20)
View(deaths_Sep6_13_20)
```

### 4. Save the csv files in Data folder for class use

```{r}
write_csv(cases_Sep6_13_20,path='Data/cases_Sep6_13_20.csv')
write_csv(deaths_Sep6_13_20,path='Data/deaths_Sep6_13_20.csv')
```